NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-06-23 paper/itemizedReviewResponse.txt

Response to the reviews

We greatly appreciate all reviewers' valuable comments. The summary of major changes is listed below. We have itemized the 44 questions/comments and answer each individually.

Summary of major changes:
- Revised section 6 to better convey the relationship among the different scenarios and provide further clarification on the result observed in 6.3.
- Revised the introduction to remove hyperbole words, clarify the NLP vs. CNN data modality comparison, and simplify the contribution list.
- Clarified the domain experts' involvement at the beginning of section 4 and link the task with the application scenarios.
- Revised section 5 to better explain the user interactions and add discussion on how to interpret the visual encodings.
- Reworked the video, included a new introduction section and added a voiceover.

@@

Q1. The usage scenarios also look interesting, especially the Scenario 3. Does the result to some extent suggest that the attention mechanism is not a very efficient component in this task? I expect the authors dig it further.

A: The domain experts indeed find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From the result from Scenario 3, we can extrapolate that the ultimate decision relies on the semantics (i.e., encodings and composed encodings in classifier). The encoder and classifier layer can swiftly adjust weights for word/composed semantics to correct prediction. But the attention layer, which only works on how individual words are aligned, affects the prediction less significantly. Recent works also support this observation in word embeddings, e.g., ELMo~\cite{salant2017contextualized}, where better word embeddings can substantially benefit a model. However, it by no means implies attention is not useful, as it serves as a way to compose word semantics although it does not directly output that. We revised section 6.3 to include this comment on the interpretation of the visualization result.

Q2. In Section 2, the neural net in NLP and attention mechanisms are introduced. But the descriptions are unsatisfactory. There is no formal definition or description of the model and the attention mechanism. It's still unclear how the attention mechanism is applied in the model. I would suggest the authors include a concise and formal introduction in the background part and include some formula.

A: The exact computation of attention can vary from model to model despite the shared overall concept, i.e., the alignment between words. As a result, when we discuss the general concept of attention we cannot fix on a single mathematical formulation. The exact formulation for the model we investigated is discussed in Appendix A.

@@

Q12, A brief guidance of how to interpret visual patterns that emerge in each view would be very beneficial. In particular, did the authors notice any systematic / repetitive patterns in the triangular prediction view? Do these patterns indicate certain issues in the model / attention / perturbation schemes?

A: As discussed in the experiment of section 6.1, the spread of the pattern directly corresponds to the stability of the prediction. A recurring pattern we noticed is illustrated in Fig. 9(d), where the given sentence pair has low stability because the original prediction is a boundary prediction (probability is in between two classes). In addition, the perturbation procedure will also affect the pattern as longer sentences will likely have more varied perturbations (contains more nouns and verbs). We include a brief discussion on the variability caused by the perturbation procedure in section 5.4.

Q13, Same thing applies to the attention heatmap, the parameter histograms, and the overview scatterplots in Figure 8 (e.g. does the curve in Figure 8d tell us something?)

A: The curve patterns in the Figure 8d are likely caused by the different number of words that was perturbed (i.e., each line pattern corresponds to sentence pairs that include the same numbers of nouns plus verbs that was perturbed). This demonstrated the added variation the perturbation procedure brings into the system. We include a brief discussion on the additional variability caused by the perturbation procedure and sentence length in section 5.3.

@@

A: these errors have been fixed.

------------------------------- Reviewer 3 -------------------------------

Q20, After the teaser image, the paper spends the first four pages on neural networks, NLP, and inference. One has to look hard to detect the lede in the next to last paragraph of the introduction ("Moreover, we propose...") Highlight it!

A: We have revised the last paragraph of the introduction to highlight the visualization contributions. The text following "Moreover, we propose..." is not the only place that describes visualization contributions, the integrated model pipeline that allows user perturbation in all stages is another major contribution, the same holds for the visual encoding for the prediction of the NLI problem.

@@

Q23, How does one handle very different sentence lengths in the sentence view?

A: One can expand the text box to include multiple lines to show longer sentences. There is no special encoding here since we are just showing the raw text.

Q24, What if the sentence structures between premise and hypothesis are not drastically different; will the matrix be as effective, and thus the graph need to be focused on?

A: If the sentence structures are not drastically different, the matrix view will still be effective which often highlight alignment as entries in the matrix diagonal. The most significant benefit of the graph is the ability for the domain expert to see the connection between words directly. In the matrix, one needs to first identify the high-value entry and then find the row and column then identify the corresponding words. If there are multiple correspondences, the user needs to check them one by one in the matrix view.

@@

Q33, Part of the problem is that the language is very dense in this paper. Although the grammar is acceptable, it was a slow read and there were some passages that I found all but indecipherable, such as the figure 7 caption, the last paragraph of section 5.5, and the bullet list at the end of the introduction.

A: We've revised the text in the mentioned sections to improve the readability.

Q34, There were many small errors that need correcting, too many to bother listing here. There was also some hyperbole ("stunning", "unprecedented", "vision", "paradigm") that distracts me as a serious reader and should be replaced. A thorough copyediting should fix both issues.

A: We removed/replaced the words mentioned by the reviewer. For example, we updated the first sentence to "Demonstrated by many recent successes, neural networks based machine learning approaches have garnered increasing popularity, and have been adopted in a wide variety of applications.". The revised paper also went through a copy-editing pass by an editor.

@@

Q35, The application scenarios in section 6 look useful taken individually, but I wonder about how they are used in combination. The scenarios essentially describe individual tasks that experts might perform in individual views. Scenario 4 goes a little further, but is brief. The paper would very much benefit from a more complete example story of how an analyst might work with the various views in concert to perform higher-level activities. For an application paper like this, some sense of those activities, and how well the tool design works for them, is needed. The last couple paragraphs of section 7 offers only a glimpse of "interrogation" as the authors say.
Their conclusion (participants believe it convenient) isn't sufficiently supported. As it stands, there isn't clear justification of the many visualization design choices. A longer scenario encompassing the larger process would go a long way toward rectifying that.

A: These different scenarios are interconnected and often arise naturally during the exploration sessions. The scenarios 1 is often used as the first step to identify the sentence pair of interests (the user can also manually type the one they interested in scenarios 5). Then the user can close examine the individual pair as shown in scenarios 2, 3, 4 (depend on what question they want to ask). We included a brief discussion on how these scenarios related to each other and expanded the scenarios 3 to include domain experts’ interpretation of the result. We didn't follow a specific exploration path of a single user, because they (from our case study) often focus on only one or two scenarios and repeatedly applied the same analysis on different data, which may not be the best way to showcase the variety of the functionalities this tool provides.

@@

A: We have revised the video and included voice over and a simple introduction at the beginning.

------------------------------- Reviewer 4 -------------------------------

Q37, I was confused by the distinction between image networks and language networks. CNNs are not only used to make binary classifications, and there are certainly challenges in image networks that do not provide "clear meaning" and seem like they would also benefit from the proposed work. In other words, I understand the focus on NLP tasks, but the reasoning for it doesn't seem accurate (in the 3rd paragraph of the introduction).

A: In the introduction, we want to highlight the input modality difference. Continuous for the image, and discrete for natural language. The difference has introduced additional challenge especially regarding optimization as well as interpretation. In addition, in CNN we often can reveal human interpretable neuron filter (but these methods cannot be easily generalized for NLP), whereas the attention is often the only interpretable part of the NLP model. We have revised the description in the introduction to better convey this message.

@@

Q38, It was unclear to me if the idea of perturbing a model is novel, or an extension/visualization of previously introduced work in this area applied to NLP NNs. If it's the former (as it seems to be), then I think this needs to be emphasized in the introduction.

A: The perturbation of nouns and verbs on itself is not novel from NLP perspective. And truly solving the challenge of producing grammatically correct and meaningful perturbation/paraphrasing is still an open research area. In our case, we believe the novelty comes from the integration of the model with the three-stage perturbation scheme in an explorable visual analytic system. We have revised the introduction to highlight the core contribution.

Q39, I appreciated the task analysis in section 4. However, it seems a bit too brief, and the tasks don't seem to directly follow from the description of the authors experience with the NLP researchers. The scenarios described in Section 6 provide much more detail about these tasks, and Section 5 clarifies them as well, but a bit more detail here would be useful. For example, it's not clear who the NLP researchers are, how many of them were interviewed, what type of NLP projects they were working on, etc.

A: We moved the description of the collaboration into the beginning of section 4, in which we described the number of NLP experts, frequency, and the duration of the meetings for this specific project. We also added a reference to application scenarios that corresponds to the tasks to help instantiate the more abstract concepts.

@@

A: We've cited this work in Section 5.2 when discussing the due visual representations.

Q41, Placing the grammar dependency tree over the matrix / bipartite graph is interesting (described in Section 5 and Section 6.4). I wonder if it would make sense to use other types of annotations here, such as those created from event extraction tools or information extraction tools that would highlight other structures (besides grammar/POS) that could be useful in helping understanding the inference model. Work by Forbes et al. supports the use of semantic hypergraphs to provide richer textual annotations alongside text sequences ("Text annotation graphs: Annotating complex natural language phenomena" from LREC 2018).

A: After discussing with domain experts, we believe the reviewer's comment would be a very interesting direction for future works. Coming up with efficient visual encoding for these types of information that go beyond grammar tree is a challenge on its own. For example, the extract structure can be a graph (instead of a tree), which likely requires a complete redesign of the existing visualization.

Q42, In Sections 6.1 and 6.2, the application scenario gives an example of having a user perturbing the model by changing the input words (switching "pile" with "heap"). This updates the system and changes the inference output. Are the successful results of such an interaction remembered by the system? That is, could NLIZE be used to tweak/update the model so that, say, the ambiguity between pile and heap is automatically and more accurately resolved in future queries?

### 2018-06-23 paper/itemizedReviewResponse.txt

Summary of major changes:
- Revised section 6 to better convey the relationship among the different scenarios and provide further clarification on the result observed in 6.3.
- Revised the introduction to remove hyperbole words, clarify the NLP vs. CNN data modality comparison, and simplify the contribution list.
- Clarified the domain experts' involvement at the beginning of section 4 and link the task with the application scenarios.
- Revised section 5 to better explain the user interactions and add discussion on how to interpret the visual encodings.
- Reworked the video, included a new introduction section and added a voiceover.

------------------------------- Reviewer 1 -------------------------------

Q1. The usage scenarios also look interesting, especially the Scenario 3. Does the result to some extent suggest that the attention mechanism is not a very efficient component in this task? I expect the authors dig it further.

A: The domain experts indeed find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From the result from Scenario 3, we can extrapolate that the ultimate decision relies on the semantics (i.e., encodings and composed encodings in classifier). The encoder and classifier layer can swiftly adjust weights for word/composed semantics to correct prediction. But the attention layer, which only works on how individual words are aligned, affects the prediction less significantly. Recent works also support this observation in word embeddings, e.g., ELMo by Salant et al., where better word embeddings can substantially benefit a model. However, it by no means implies attention is not useful, as it serves as a way to compose word semantics although it does not directly output that. We revised section 6.3 to include this comment on the interpretation of the visualization result.

@@

Q16, A brief discussion of the applicability to other languages would be interesting maybe? Any anticipated challenges?

A: Since we have utilized Unicode text encoding throughout the code base, the proposed system can be easily adapting to other alphabetical languages (however, we have not tested the capability). One potential challenge may be originated from NLP training dataset, as there are not many large-scale training data in languages other than English.

Q17, Add label to the color scale in Figure 3 (the caption mentions attention by just to avoid confusion). Same for Figure 7a (frequency to the vertical axis), Figure 8b, and Figure 9a-c.

A: We've added "attention" label to the color scale and the "frequency" label to the histograms.

### 2018-06-26 paper/intro.tex
grammar fix

% (What is the high level motivation)
% What is the obstacle for effectively apply machine learning model
%Based on some stunning recent advances, neural networks based machine learning approaches have gained unprecedented popularity, and have been adopted in a wide variety of applications.
Demonstrated by many recent successes, neural networks based machine learning approaches have garnered increasing popularity, and have been adopted in a wide variety of applications.
%
However, researchers as well as practitioners often need to overcome many obstacles during the training, debugging, and tuning processes to realize the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made is critical to both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and propose hypotheses for the cause of the failure is key to identifying the limitations of and improving upon existing models.
However, providing meaningful answers to these questions is challenging and has been described as impossible by many.

@@

% Beside explore the internal states of the model, understand how different parts of the model work together and how one may affect one another is as or if not more important.
Despite a number of recent advances, the majority of existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory nature of the model interpretation often leads to many \emph{``what if ...''} types of questions,
such as, what if we perturbed the current input?
Will the prediction be stable?
What if we change one of the critical internal states of the model?
How would the modification affect the prediction?
What if the current prediction is wrong?
How and where could we apply minimal change to the model to produce the correct result? And how would it affect the internal state we care about? These types of queries form a natural way to gain understanding and develop hypotheses of the model mechanisms by \emph{interrogating} how the components of a model interact with each other in a dynamic setting.
Hypothetically, we can code specialized experiments for each of these scenarios.
However, such a process is not only tedious but also ignores the iterative nature of the exploratory analysis. Often, these questions are not pre-determined.
Instead, new exploration paths arise as we investigate and analyze previous observations.

Here, we aim to provide immediate and informative answers to these \emph{``what if''} questions by combining the expressive power of visualization and direct online query/optimization of the neural network model. Instead of viewing the models as invariant objects, we approach the interpretability challenge by studying them in a dynamic environment. By employing a perturbation-driven scheme, we probe the internal states of the model and examine how changes in one part of the pipeline (the input, internal states, and output prediction, see Fig.~\ref{fig:modelPipeline}) affect others, which in turn provides a new perspective to address the model interpretation problem.

We materialize the goal of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
%We implement our vision of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
However, the components of the visualization and the overall concept can be readily extended to other NLP tasks, such as question and answer, text summarization, etc.
%
In the simplest form, the inference task asks whether the
relationship between \textbf{sentence A} and \textbf{sentence B}: (1) is \emph{entailment} (one can infer \textbf{B} from \textbf{A}), (2) \emph{contradiction} (\textbf{B} disagrees with \textbf{A}) or (3) is \emph{neutral} (\textbf{A} and \textbf{B} talk about different/unrelated things).
% we can infer the content of \textbf{sentence B} from \textbf{sentence A} (\emph{entail}), if they contradict with each other (\emph{contradiction}), or whether two sentences talk about different/unrelated things (\emph{neutral}).
Natural language inference addresses the fundamental challenge of identifying semantic relationships between sentences, and is a core NLP task (see Section~\ref{sec:languageInference} for details).

One recent advance in neural natural language process models is the
introduction of \emph{attention
  mechanisms}~\cite{bahdanau2014neural,VaswaniShazeerParmar2017}
(Section~\ref{sec:attention}). Intuitively, attention asks which parts of the
input are deemed more important for making a prediction. The attention is often represented via weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models.
The proposed tool introduces a perturbation-driven visual analytics enviroment, where the domain experts can study how changes in sentence input, attention, or prediction affect each other, which helps them develop deeper intuitive and alternative hypotheses.
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label.
%Another word,
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
In addition, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.

>>>>

% (What is the high level motivation)
% What is the obstacle for effectively apply machine learning model
%Based on some stunning recent advances, neural networks based machine learning approaches have gained unprecedented popularity, and have been adopted in a wide variety of applications.
As demonstrated by many recent successes, neural network-based machine learning approaches have garnered increasing popularity and have been adopted in a wide variety of applications.
%
However, researchers as well as practitioners often need to overcome many obstacles during the training, debugging, and tuning processes to realize the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical for both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and propose hypotheses for the cause of the failure is key to identifying the limitations of and improving upon existing models.
However, providing meaningful answers to these questions is challenging and has been described as impossible by many.

@@

% Beside explore the internal states of the model, understand how different parts of the model work together and how one may affect one another is as or if not more important.
Despite a number of recent advances, most of existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory nature of the model interpretation often leads to many \emph{``what if ...''} types of questions,
such as, what if we perturbed the current input?
Will the prediction be stable?
What if we change one of the critical internal states of the model?
How would the modification affect the prediction?
What if the current prediction is wrong?
How and where could we apply minimal change to the model to produce the correct result? And how would it affect the internal state we care about? These types of queries form a natural way to gain an understanding and develop hypotheses of the model mechanisms by \emph{interrogating} how the components of a model interact with each other in a dynamic setting.
Hypothetically, we can code specialized experiments for each of these scenarios.
However, such a process is not only tedious but also ignores the iterative nature of the exploratory analysis. Often, these questions are not pre-determined.
Instead, new exploration paths arise as we investigate and analyze previous observations.

Here, we aim to provide immediate and informative answers to these \emph{``what if''} questions by combining the expressive power of visualization and the direct online query/optimization of the neural network model. Instead of viewing the models as invariant objects, we approach the interpretability challenge by studying them in a dynamic environment. By employing a perturbation-driven scheme, we probe the internal states of the model and examine how changes in one part of the pipeline (the input, internal states, and output prediction, see Fig.~\ref{fig:modelPipeline}) affect others, which in turn provides a new perspective to address the model interpretation problem.

We materialize the goal of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
%We implement our vision of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
However, the components of the visualization and the overall concept can be readily extended to other NLP tasks, such as question and answer, text summarization, etc.
%
In its simplest form, the inference task asks whether the
relationship between \textbf{sentence A} and \textbf{sentence B}: is (1) \emph{entailment} (one can infer \textbf{B} from \textbf{A}), (2) \emph{contradiction} (\textbf{B} disagrees with \textbf{A}) or (3) \emph{neutral} (\textbf{A} and \textbf{B} talk about different/unrelated things).
% we can infer the content of \textbf{sentence B} from \textbf{sentence A} (\emph{entail}), if they contradict with each other (\emph{contradiction}), or whether two sentences talk about different/unrelated things (\emph{neutral}).
Natural language inference addresses the fundamental challenge of identifying semantic relationships between sentences and is a core NLP task (see Section~\ref{sec:languageInference} for details).

One recent advance in neural natural language process models is the
introduction of \emph{attention
  mechanisms}~\cite{bahdanau2014neural,VaswaniShazeerParmar2017}
(Section~\ref{sec:attention}). Intuitively, attention asks which parts of the
input are deemed more important for making a prediction. Attention is often represented via weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many theories about how attention works in various models.
The proposed tool introduces a perturbation-driven visual analytics environment, where the domain experts can study how changes in sentence input, attention, or prediction affect each other, which helps the experts develop deeper intuitive and alternative hypotheses.
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label.
%Another word,
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
In addition, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.
